---
title: "Homework 10"
author: "Stephen Peters-Collaer"
date: "3/30/2022"
output: html_document
---

Click [here](index.html) to return to the main page

I worked next to Lindsey Carlson on this assignment, and we discussed some of the questions together.

**Question 1:**
Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.
```{r Question 1}
##########################
# FUNCTION:zero_count
# purpose: Count the nuber of zeros in a numeric vector
# input:A numeric vector
# output: A single number that is a count of the number of zeros in that vector
# ------------------------

zero_count <- function(vector = sample(-100:100, 1000, replace = TRUE)){
  counter <- 0
for (i in 1:length(vector)) {
  if(vector[i] == 0){ 
    counter <- counter+1
  } # end if statement
} # end forloop
    return(counter)
} # end function

  zero_count()

```

**Question 2: **
Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r, Question 2}
##########################
# FUNCTION:zero_count_short
# purpose: Count the nuber of zeros in a numeric vector using subsetting
# input:A numeric vector
# output: A single number that is a count of the number of zeros in that vector
# ------------------------
zero_count_simple <- function(vector = sample(-100:100, 1000, replace = TRUE)){
  return(length(vector[vector==0]))
}
zero_count_simple()



```